Honey Bee Colonies Impacted by Varroa, American Foulbrood and Global Warming

### 2024-08-01 starter-analysis-exercise/products/report/manuscript_final-2.qmd
Updated HTML, photos, and references included with citations

---
title: "Honey Bee Colonies Impacted by Varroa, American Fouldbrood and Global Warming"
subtitle: ""
author: Edward Cruz, Jr, Leonel G. Salazar, Amelia Tristan, Mojisola Popoola
date: today
format:
  html:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../citations.bib
csl: ../apa.csl
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
library(ggplot2)
library(dplyr)
library(scales)
library(showtext)
library(ggimage)
```

@@

Group Ten is conducting a comprehensive analysis of historical data from
various public agencies to evaluate the impact of Varroa mites, American
Foulbrood, and global warming on hive losses in the United States.This
research utilizes extensive datasets from the National Agricultural
Statistics Service, the Agricultural Statistics Board, and the United
States Department of Agriculture (USDA), encompassing multiple years of
data. The data highlights the topic on hive losses attributed to mites,
bacterial infections, and environmental factors related to global
warming. Through advanced data visualization techniques in R, we aim to
demonstrate and validate the detrimental effects of these factors on
honeybee colonies, highlighting the consequent implications for honey
production and broader food security. \[\@hivecollapse\]

@@

Honeybees have served as nature's pollinators for centuries. With their
relationship with humans documented as far back as ancient Egyptian and
Hindu cultures. Historically, humans have maintained beehives and
utilized honey for its medicinal properties in various civilizations,
including the Egyptians, Assyrians, Chinese, Greeks, and Romans. Honey's
natural antibacterial qualities made it a valuable treatment for wounds.
This was a practice continued by Romans and Russians during World War I.
Honeybees and other pollinators are essential for food production and
nutritional security. But even though they face numerous survival
challenges Varroa mites pose a significant threat to bee colonies. These
tiny red-brown parasites live on adult honeybees and reproduce on larvae
and pupae. Another major threat is American Foulbrood Disease (AFB).
This is caused by the bacterium Paenibacillus larvae, which is fatal to
honeybee larvae and requires incineration of infected hives to prevent
its spread. The impact of climate change on honeybee colony losses is a
growing area of research. While correlations have been found between
higher winter temperatures and increased colony losses, the effects of
warmer autumn and winter temperatures on colony dynamics and survival
require further investigation.

@@

@@

Written By:sdns6mchl4. (2016, February 24). Varroa mite spread in the
United States. Beesource Beekeeping Forums.
https://www.beesource.com/threads/varroa-mite-spread-in-the-united-states.365462/
\## General Background Information

@@

![](/images/varroa-mite-bee-01.webp){fig-align="center" width="565"}

@@

#Citations

>>>>

---
title: "Honey Bee Colonies Impacted by Varroa, American Foulbrood and Global Warming"
subtitle: ""
author: Edward Cruz, Jr., Leonel G. Salazar, Amelia Tristan, Mojisola Popoola
date: today
format:
  html:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../citations.bib 
csl: ../apa.csl
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
library(ggplot2)
library(dplyr)
library(scales)
library(showtext)
library(ggimage)

@@

Group Ten is conducting a comprehensive analysis of historical data from
various public agencies to evaluate the impact of Varroa mites, American
Foulbrood, and global warming on hive losses in the United States.This
research utilizes extensive datasets from the National Agricultural
Statistics Service, the Agricultural Statistics Board, and the United
States Department of Agriculture (USDA), encompassing multiple years of
data. The data highlights the topic on hive losses attributed to mites,
bacterial infections, and environmental factors related to global
warming. Through advanced data visualization techniques in R, we aim to
demonstrate and validate the detrimental effects of these factors on
honeybee colonies, highlighting the consequent implications for honey
production and broader food security. ***Dennis and Kemp's study on
honey bee hive collapse [@hivecollapse] provides important insights into
Allee effects and ecological resilience.***

@@

Honeybees have served as nature's pollinators for centuries. With their
relationship with humans documented as far back as ancient Egyptian and
Hindu cultures. Historically, humans have maintained beehives and
utilized honey for its medicinal properties in various civilizations,
including the Egyptians, Assyrians, Chinese, Greeks, and Romans. Honey's
natural antibacterial qualities made it a valuable treatment for wounds.
This was a practice continued by Romans and Russians during World War I.
Honeybees and other pollinators are essential for food production and
nutritional security. But even though they face numerous survival
challenges Varroa mites pose a significant threat to bee colonies. These
tiny red-brown parasites live on adult honeybees and reproduce on larvae
and pupae. Another major threat is American Foulbrood Disease (AFB).
This is caused by the bacterium Paenibacillus larvae, which is fatal to
honeybee larvae and requires incineration of infected hives to prevent
its spread. The impact of climate change on honeybee colony losses is a
growing area of research. While correlations have been found between
higher winter temperatures and increased colony losses, the effects of
warmer autumn and winter temperatures on colony dynamics and survival
require further investigation. ***Dennis and Kemp's study on honey bee
hive collapse [@hivecollapse] provides important insights into Allee
effects and ecological resilience.***[@usda_honey_bees; @usda_nass]

@@

***[@varroamap] Varroa Map of infestation periods***

@@

[@epa_climate_indicator] ***Global warming trends from 1900-2024***

***Spread of Varroa Mite by Year of First Apperance [@varroamap]***
Written By:sdns6mchl4. (2016, February 24). Varroa mite spread in the
United States. Beesource Beekeeping Forums.
https://www.beesource.com/threads/varroa-mite-spread-in-the-united-states.365462/

\## General Background Information

@@

![](images/varroa-mite-bee.jpg){fig-align="center"}

@@

##The full analysis is still pending###
